Marine Conservation In Fishing Effort Markets

### 2019-05-22 docs/old_ms/PIPA_Palau_Science.tex
Updated MS for John and Chris

\documentclass[9p,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\usepackage{placeins}
\usepackage{comment}

\title{Large-Scale Marine Protected Areas in the World's Largest Tuna Fishery}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Juan CarlosVillase\~{n}or-Derbez}
\author[b]{John Lynham}

\affil[a]{Bren School of Environmental Science \& Management, University of California at Santa Barbara, Santa Barbara, CA}
\affil[b]{Department of Economics, University of Hawaii at Manoa, Honolulu, HI}

% Please give the surname of the lead author for the running footer
\leadauthor{Villase\~{n}or-Derbez}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{The oceans are becoming increasingly crowded, with different activities competing for space. Marine Protected Areas are inherently spatial, and have the primary objective of conserving bounded waters by displacing or limiting fishing effort. Our work shows that when a fishery is managed by selling effort rights (\emph{e.g.} with a Vessel-Day Scheme), spatial closures displace fishing effort at a cost to the implementing country, with benefits to other countries. Costs can be offset if effort rights are tradable across countries. With global conservation targets seeking to protect 10\% of the world’s ocean by 2020, it is important that the spatial displacement of economic activities is fully considered.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{All authors contributed equally to this work}
\authordeclaration{The authors declare that they have no conflicts of interest}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: juancarlos@ucsb.edu}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Marine Spatial Planning $|$ Fisheries $|$ Marine Conservation}

\begin{abstract}
The Parties to the Nauru Agreement (PNA) is a system for managing the world’s largest skipjack tuna fishery, covering an area of 14.6 million km\textsuperscript{2} in the Pacific Ocean. The fishery in the region operates under a Vessel-Day Scheme (VDS), which sells access rights that allow vessels to fish in PNA waters. In 2015, PNA member Kiribati implemented one of the world's largest conservation areas: the Phoenix Islands Protected Area (PIPA, 397,447 km\textsuperscript{2}), effectively excluding all tuna purse seining activities. Such an intervention is likely to have effects on vessel spatial distribution and behavior, as well as short-term costs (due to the reduction in fishable area). We use identification of fishing activity via Automatic Identification Systems and causal inference techniques to evaluate the effect of PIPA on vessel distribution and behavior, as well as costs to Kiribati and the PNA. We find a crowding effect within PNA waters after the implementation of the protected area. In the first year of the closure, there is no drop in total fishing effort within Kiribati's EEZ and a reported increase in revenue from access rights sold. However, from 2016 onward there is a noticeable drop in fishing effort within Kiribati and a reported drop in license and access fee revenue. These changes are robust to controlling for the effects of El Ni\~{n}o on spatial distribution of fishing effort. At the same time, fishing effort increases in other parts of the PNA. This redistribution of fishing effort eventually results in a reduction of 5,195 vessel-days in Kiribati, which implies a loss of around \$50 million USD; but much lower losses are reported in official government reports (\$30.5 million USD), suggesting that the tradable nature of the VDS system is buffering the cost for Kiribati. We use our results to inform predictions of the impacts of a proposed Large-Scale Marine Protected Area (LSMPA) in Palau (another PNA member) and estimate potential losses to range from \$2.5 to \$11 million annually. PNA members who indirectly benefit from MPAs should consider mechanisms that reward such conservation actions.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

@@

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

@@

\dropcap{H}umans are increasingly utilizing the oceans. Multiple ocean uses such as off-shore aquaculture, conservation, energy harvesting, deep-sea mining, and fisheries are likely to compete for space. As we move forward with blue growth, we must understand the potential effects of activities displacing each other and establish causal links between past management interventions and their outcomes \citep{burgess_2018}. One of the most notable spatial interventions is the creation of no-take Marine Protected Areas (MPAs), which seek to conserve the environment by eliminating fishing effort within their waters.

Global international goals aim to protect 10\% of the ocean environment by 2020. In an effort to meet this target, there has been a rapid increase in MPA coverage \citep{wood_2008,sala_2018}, largely driven by a small number of Large-Scale Marine Protected Areas (LSMPAs; areas larger than 30,000 km\textsuperscript{2} \emph{sensu} \cite{desanto_2013}). Today, a small number of LSMPAs represent at least 80\% of the protected area in the ocean (Fig. \ref{fig:LSMPAs_map}; \citep{toonen_2013}). However, very little is known about their human dimensions and implications for fisheries \citep{gray_2017}. Furthermore, most research on LSMPAs has focused on their potential ecological benefits, but have left aside the economic implications. One issue of particular importance is that of the displacement or redistribution of fishing effort, which may influence the outcomes of a spatial closure and represent large opportunity costs \citep{smith_2003,smith_2010}.

\begin{figure}[H]
\centering
\includegraphics{img/LSMPAs_map.pdf}
\caption{\label{fig:LSMPAs_map}Large Scale Marine Protected Areas. The map shows all areas larger than 30,000 Km\textsuperscript{2}. The Phoenix Islands Protected Area is shown in red.}
\end{figure}

The Phoenix Islands Protected Area (PIPA) in Kiribati is one of the most notable Large-Scale Marine Protected Areas on earth. Implemented on January 1st of 2015, PIPA closed an area of 397,447 km\textsuperscript{2} to fishing and was implemented within an area where approximately 50\% of the world's tuna is caught. Tuna purse seine fisheries in the region are collectively managed under a Vessel-Day Scheme (VDS) by nine countries commonly referred to as the Parties to the Nauru Agreement (PNA). Members include the Federated States of Micronesia, Kiribati, the Marshall Islands, Nauru, Palau, Papua New Guinea, the Solomon Islands, and Tuvalu; Tokelau joined the VDS in 2012 and started selling access rights in 2013 (Figure \ref{fig:PNA_map}). The Nauru Agreement regulates access of foreign vessels (\emph{i.e.} those from non-PNA countries). Holding 80\% (14.6 million km\textsuperscript{2}) of historical skipjack tuna purse seine grounds within their Exclusive Economic Zones (EEZ), PNA countries have achieved greater bargaining power when providing fishing access to foreign fleets \cite{havice_2010}. The vessel-day price rose from \$5,000 USD in 2012 to at least \$9,000 USD in 2016, driven by increasing price floors. The revenue from access fees may represent up to 50\% of government revenue for some of the members.

A spatial closure the size of PIPA is likely to cause changes in the spatial distribution and behavior of fishing vessels. For example, the anticipation of LSMPAs may lead to preemptive overfishing, which likely erodes or delays the expected benefits of an intervention \cite{mcdermott_2018,hanich2018unraveling}. Under a VDS, a reduction in total fishing area within one country's EEZ will result in a reduction in license revenues to said country. However, the benefits of the spatial closure are dispersed amongst all other PNA members (through fish movement), who in turn benefit from the conservation efforts of the initial country. While no studies have assessed the impact of PIPA, other PNA members have pledged the implementation of LSMPAs by 2020 (\emph{i.e.} Palau).

We simulate the PNA fishery with the addition of spatial closures to characterize possible outcomes of such interventions. We then empirically evaluate the behavioral responses and spatial redistribution of the industrial tuna purse seine fleet resulting from the implementation of the Phoenix Islands Protected Area, and quantify its economic ramifications and impacts to Kiribati. We use the same data to hypothesize what might be the impacts of the proposed Palau National Marine Sanctuary. These are two of the largest protected areas on the planet and both are controlled by PNA countries.

Our empirical approach uses identification of fishing activity via Automatic Identification Systems (AIS) to track 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018. We continuously observe 92 vessels for the 2012 - 2018 period. Of these, 64 vessels fished within PIPA at least once prior to its implementation, and we refer to them as the ``displaced vessels''. The remaining 28 vessels never fished in PIPA waters, and we refer to them as ``non-displaced vessels''. The group with the remaining 221 vessels contains vessels that were not continuously observed before and after the implementation of PIPA, and we refer to these as ``other vessels''.

\begin{figure}[htbp]
\centering
\includegraphics{img/PNA_model.pdf}
\caption{\label{fig:PNA_model}Change in revenue from the PNA model with conservation. Revenue is shown for Kiribati alone (A) and all other PNA countries (B) as a function of increasing proportion of patch as reserve ($R$). Each line (color) represents a value of the movement parameter $\theta$.}
\end{figure}

@@

\subsection{PNA model with conservation}

We simulate the PNA as a ten-patch meta-population system with discrete time, where Patch 1 considers the implementation of a spatial closure. Patches 2-9 are the other PNA countries, and Patch 10 represents the high seas and non-PNA countries. The stock remains within a Patch during the season, but escapement (\emph{i.e.} stock minus catch) is distributed between all countries at the end of each year. A detailed explanation of the model is presented in the Methods section. We find that a spatial closure in Patch 1 always results in a loss or no-change in revenue from vessel-days, even when the stock moves freely between the protected and non-protected portion of the Patch. The loss in revenue increases with reserve size, but decreases as within-patch movement increases (Fig. \ref{fig:PNA_model}A). For all other PNA countries, however, a spatial closure in Patch 1 is beneficial, especially as within-patch movement decreases and reserve size increases (Fig. \ref{fig:PNA_model}B). The inverse effects of the spatial closure on Kiribati and the other PNA countries are driven by the redistribution of escapement. The stock not fished in the protected portion of Patch 1 eventually redistribute to the other patches, which increases stock size and causes vessel-day prices to increase. This model shows that the costs of conservation are incurred by Patch 1, but the benefits are perceived by the eight other patches. Moreover, the gains in revenue for other PNA countries does not compensate for the total losses to Kiribati, as a large portion of effort is redistributed to the High Seas.

\subsection{Empirical analysis}

Spatial distribution and longitudinal shifts of tuna -particularly skipjack- have been linked to ENSO events \cite{lehodey_1997}. Vessels may then redistribute across PNA countries as they track these stocks \cite{aqorau_2018}. To account for environmentally-driven patterns in vessel location, we incorporate the monthly NINO4 anomaly index in our analyses. The NINO4 index is an area-averaged measure of Sea Surface Temperature from 5S-5N and 160E-150W, made available by the Global Climate Observing System (Fig. \ref{fig:nino_plot}). The NINO4 anomaly index is the same index, with the 1981-2010 mean removed.

We first inspect the crowding effects that may arise due to the net reduction in fishing area. We produced 1-degree rasters of monthly fishing effort for our displaced and non-displaced vessels, and calculated two indices of spatial overlap between them: 1) the number of cells that had fishing activity from both groups for each month and 2) the correlation of presence/absence of fishing activity between both groups over one month. We find that the two fleets significantly interact more with each other after the implementation of PIPA (Table \ref{tab:sp_corr} Fig. \ref{fig:sp_corr}). The number of cells with presence from both fleets and spatial correlation increase by a factor of four and three, respectively. This increase in crowding is likely to increase the encounter rates with other vessels, and reduce the efficiency of fishing operations. This might cause vessels to leave their current fishing grounds and re-optimize their spatial effort, leading to a subsequent decrease as the crowding measures return towards pre-implementation levels. These results are robust to a series of specifications, accounting for the addition of satellites and monthly NINO4 anomaly (Table \ref{tab:sp_corr}). A model fit with the NINO4 anomaly as the only explanatory variable explained just 3\% and 7\% of the variation in our crowding measures. Similar results are obtained when repeating the exercise for Kiribati only (Table \ref{tab:KIR_sp_corr}, Fig \ref{fig:KIR_sp_corr})

\begin{figure}[ht]
\centering
\includegraphics{img/sp_corr.pdf}
\caption{\label{fig:sp_corr}Number of cells that had displaced and non-displaced vessels (A) and spatial correlation in the presence-absence of each group per cell (B). The solid lines represent the 4\textsuperscript{th} degree polynomial fit reported in \ref{tab:sp_corr}. Note that the late 2016 and early 2017 showed negative or neutral NINO4 anomalies, similar to those in the pre-PIPA period.}
\end{figure}

The behavioral responses that vessels can have to a spatial closure may occur in different ways. For example, displacement to new fishing grounds may represent a cost as fishers search the ocean to identify the most suitable fishing spots. This may result in increased fuel and labor costs. For every vessel in each group, we calculate nine key measures that could capture responses to spatial closures: daily fishing hours, daily non-fishing at-sea hours, the proportion of fishing to non-fishing hours at sea hours, daily distance traveled, daily mean distance from shore of fishing events (km), daily mean distance from port of fishing events (km), as well as monthly hours spent in PNA waters, Kiribati waters, and the High Seas (Fig. \ref{fig:all_panels}). We leverage our Before-After-Control-Impact (BACI) design and implement a log-linear difference-in-differences analysis to evaluate how these measures change for displaced vessels after implementation of PIPA, relative to the trends observed for non-displaced vessels (see the Methods section for our empirical specification). As before, our analyses incorporate monthly NINO4 anomalies to account for possible environmentally-driven variations.

We find no evidence of displaced vessels fishing for more hours after PIPA implementation, and in fact observe a negative effect (24.3\% decrease; $p < 0.01$; Table \ref{tab:main_DID}) relative to the non-displaced vessels. Likewise, we observe a 3.4\% decrease of fishing hours relative to total at-sea hours ($p < 0.01$; Table \ref{tab:main_DID}). Treated vessels traveled 21\% less distance, with fishing events occurring 28.5\% and 15.7\% closer to shore and to port, respectively. These changes in distance from shore and port are likely caused by redistribution, as we observe that displaced vessels fish 58.6\% and 40.3\% less in Kiribati and PNA waters, compared to the trend observed for non-displaced vessels ($p < 0.01$). We do not observe a statistically significant increase in fishing hours on the High Seas.

These patterns suggest that displaced vessels are fishing less overall and this decrease is driven by a decline in fishing within PNA waters (including Kiribati). These vessels are also fishing closer to their home ports. Vessels that did not use to fish in PIPA are fishing more in Kiribati and PNA waters, with slight increases in the High Seas (Figure \ref{fig:fishing_raster_diff}). The results are robust to a set of different specifications, with interaction effects shown in Fig  \ref{fig:other_specifications}. We repeat this analysis for sample restrictions where we exclude all Chinese vessels (Table \ref{tab:DID_without_CHN}), all PNA-owned vessels (Table \ref{tab:DID_without_PNA}) and all Taiwanese and USA vessels (Table \ref{tab:DID_without_USA_TWN}) and find qualitatively the same reposnses.

\subsubsection{Economic impacts}

The crowding effect combined with the reduction in hours spent in Kiribati and PNA waters overall suggests that displaced vessels have redistributed elsewhere, meaning that they buy less vessel-days from PNA countries. To quantify the potential impacts of this leakage, we estimate the total annual vessel-days received by Kiribati and all PNA countries by each group of vessels (Fig. \ref{fig:all_PS_VDS_year}), and convert these to license revenues using a conservative vessel-day price of \$9,000 USD\footnote{The Pacific Islands Forum Fisheries Agency \emph{Tuna Development Indicators 2016} report states that ``Days are currently [2016] selling in a range between \$9,000 and \$13,000 USD.''}. We look at all 313 vessels to obtain a more accurate representation of total revenues, but continue to group vessels as displaced (n = 64), non-displaced (n = 28), and other vessels (n = 221).

We find that between 2015 and 2016, displaced vessels spent 3,916 and 2,249 less vessel-days in Kiribati and PNA waters, respectively (Figs. \ref{fig:all_PS_VDS_year} and \ref{fig:included_PS_VDS_year_DiD}). Over the same period, non-displaced and other vessels spent 1,278 less days in Kiribati, but spent 9,853 more days in PNA waters overall. These changes result in a net loss of 5,195 vessel-days for Kiribati, and a net gain of 7,600 vessel-days at the PNA-level (\emph{i.e.} the other 8 countries). The net reduction of vessel-days in Kiribati represents a loss of \$46.7 million USD in vessel-day licenses, while the net gain at the PNA-level results in an increase of \$68 million USD.

Moreover, annual vessel-days in Kiribati continued to decrease to just 7,479 in 2018 (Fig. \ref{fig:all_PS_VDS_KIR_year}). This trend is mainly caused by displaced vessels allocating less time to Kiribati (Figs \ref{fig:all_PS_VDS_KIR_year} and \ref{fig:hist_kir_fishing}). Looking at the total annual vessel-days allocated by all vessels to all countries, we see that the largest reductions occur for Kiribati, while Papua New Guinea exhibits a proportional increase (Fig. \ref{fig:all_PS_VDS_cty_year}).

\begin{figure}[ht]
\centering
	\includegraphics{img/all_PS_VDS_year.pdf}
	\caption{\label{fig:all_PS_VDS_year}Observed vessel-days for all PNA countries by displaced, non-displaced, and other vessels.}
\end{figure}

\begin{figure}[ht]
\centering
	\includegraphics{img/included_PS_VDS_year_DiD.pdf}
	\caption{\label{fig:included_PS_VDS_year_DiD}Vessel days spent inside A) PNA waters and B) Kiribati waters by vessel group.}
\end{figure}

We complement our analysis of change in observed vessel-days by looking at country-level data. Specifically, we use data compiled by the Pacific Islands Forum Fisheries Agency (FFA\footnote{https://www.ffa.int/node/2050}) where annual revenues from license fees (for VDS and other access programs) are reported for each country (2008 - 2016; Fig. \ref{fig:revenues}A). We find that Kiribati's revenue went from \$148.8 million USD in 2015 to \$118.3 million USD in 2016, representing a decrease of \$30.5 million USD. However, total PNA revenues showed a net increase of \$28 million USD (Fig. \ref{fig:total_PNA_revenues}). The largest decrease was observed for Kiribati, while the largest increase was observed for Papua New Guinea. This is suggestive but certainly not conclusive evidence that Kiribati may have been able to lessen the impact of PIPA on their license revenues by selling their fishing days to other members of the PNA (including Papua New Guinea).

\begin{figure}[ht]
\centering
	\includegraphics{img/revenues.pdf}
	\caption{\label{fig:revenues}
		License revenue for PNA countries. A) Annual revenue from fishing license fees by country and year (2008 - 2016) B) $log_{10}$-transformed FFA-reported revenues vs. the revenues inferred from vessel activity observations (2012 - 2016). The dashed line represents line of best fit, solid line represents 1:1 line. The same graph using absolute values is shown in \ref{fig:revenue_FFA_GFW_linear}.}
\end{figure}

%Need to change the above to Skipjack catch at some point

>>>>

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[12pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% http://www.sciencemag.org/authors/preparing-manuscripts-using-latex 
% This package should properly format in-text
% reference calls and reference-list numbers.

\usepackage{scicite}

\usepackage{times}

% Packages I manually added
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{float}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{pdflscape}

% The preamble here sets up a lot of new/revised commands and
% environments.  It's annoying, but please do *not* try to strip these
% out into a separate .sty file (which could lead to the loss of some
% information when we convert the file to other formats).  Instead, keep
% them in the preamble of your main LaTeX source file.

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% Include your paper's title here

\title{Marine Conservation In Fishing Effort Markets}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author{Juan CarlosVillase\~{n}or-Derbez,$^{1\ast}$ John Lynham,$^{2}$ Christopher Costello$^{1}$\\
\\
\normalsize{$^{1}$Bren School of Environmental Science \& Management,}\\
\normalsize{University of California at Santa Barbara, Santa Barbara, CA}\\
\normalsize{$^{2}$Department of Economics, University of Hawaii at Manoa, Honolulu, HI}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail: juancarlos@ucsb.edu.}
}

% Include the date command, but leave its argument blank.

\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

@@

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}
Our oceans are becoming increasingly crowded and privatized as different activities compete for space. Marine Protected Areas have the primary objective of conserving bounded waters by displacing or limiting fishing effort. With global conservation targets seeking to protect 10\% of the world’s ocean by 2020, it is important that the spatial displacement of economic activities is fully considered. A spatial closure in a vessel-day-managed fishery can potentially result in large costs to the implementing country, which might deter conservation efforts. We show that effort markets may reduce up to 70\% of the costs, and that allocation rules based on biomass (rather than effort) may incentivize conservation. We then use identification of fishing activity via Automatic Identification Systems to evaluate the effect that the Phoenix Islands Protected Area (PIPA, 397,447 km\textsuperscript{2}, implemented implemented in 2015 in Kiribati) had on vessel distribution and behavior. Aggregate fishing effort remains relatively constant following the closure, but is significantly reduced within Kiribati, especially by vessels that historically fished within PIPA before its implementation. We find evidence that Kiribati may be trading its surplus of vessel-days, which would significantly reduce the costs of displacing fishing effort. We use these insights predict the implications of a proposed Large-Scale Marine Protected Area (LSMPA) in Palau (another PNA member) and estimate potential losses to range from \$2.5 to \$11 million annually. Our work highlights the role of effort markets in enabling conservation, and the importance of allocation rules that praise, not punish, conservation. The Parties to the Nauru Agreement have successfully managed their fishery and developed a management scheme that allows for conservation interventions to come at a low cost.
\end{sciabstract}

@@

Humans are increasingly utilizing the oceans. Multiple ocean uses such as off-shore aquaculture, conservation, energy harvesting, deep-sea mining, and fisheries are likely to compete for space. As we move forward with blue growth, we must understand the potential effects of activities displacing each other and establish causal links between past management interventions and their outcomes \cite{burgess_2018}. One of the most notable spatial interventions is the creation of no-take Marine Protected Areas (MPAs), which seek to conserve the environment by eliminating fishing effort within their waters.

Today, a small number of Large-scale Marine Protected Areas (LSMPAs; areas larger than 30,000 km\textsuperscript{2} \emph{sensu} \cite{desanto_2013}) represent at least 80\% of the protected area in the ocean (Fig. \ref{fig:LSMPAs_map}; \cite{toonen_2013}). However, very little is known about their human dimensions and implications for fisheries \cite{gray_2017}. Furthermore, research on LSMPAs has focused on their potential ecological benefits, but have left aside the economic implications. One issue of particular importance is that of the displacement or redistribution of fishing effort, which may influence the outcomes of a spatial closure and represent large opportunity costs \cite{smith_2003,smith_2010}.

Displacement or elimination of fishing effort within the protected area might come at a cost when fisheries are managed under an effort market. For example, tuna purse seine fisheries in the Pacific are collectively managed under a Vessel-Day Scheme (VDS) by nine countries commonly referred to as the Parties to the Nauru Agreement (PNA; Figure \ref{fig:PNA_map}). The agreement regulates the access of foreign fishing vessels to PNA waters, and control the allocation and sales of 45,000 vessel-days. Holding 80\% of historical skipjack tuna purse seine grounds within their joint Exclusive Economic Zones (EEZ; with surface area of  14.6 million km\textsuperscript{2}), PNA countries have achieved greater bargaining power \cite{havice_2010}; vessel-day price rose from \$5,000 USD in 2012 to at least \$9,000 USD in 2016, driven by increasing price floors. The revenue from access fees may represent up to 50\% of government revenue for some of the parties.

The Phoenix Islands Protected Area (PIPA) in Kiribati is one of the most notable Large-Scale Marine Protected Areas on earth. Implemented on January 1st of 2015, PIPA closed an area of 397,447 km\textsuperscript{2}, about 11\% of Kiribati's EEZ or 2.7\% of total PNA area, effectively displacing all fishing effort within it \cite{mcdermott_2018,mccauley_2016}. A pressing question is that of the costs associated to the displacement of fishing effort and possible revenue losses from the fishing license sales. While no studies have assessed the implications of PIPA, other PNA members have pledged the implementation of LSMPAs, like Palau's 80\% closure to go into effect in 2020 \cite{cimino_2019}.

We develop a bioeconomic model of a fishery operating under a vessel-day scheme with spatial closures to explore the effects of effort redistribution resulting from MPA implementation. We characterize the likely outcomes with and without vessel-day trading between countries, as well as the effect of different allocation rules. We then empirically evaluate the behavioral responses and spatial redistribution of the industrial tuna purse seine fleet operating within PNA waters after the implementation of the Phoenix Islands Protected Area, and quantify its economic ramifications and implications to Kiribati. We use the same data and insights from our model to hypothesize what might be the impacts of the proposed Palau National Marine Sanctuary. These are two of the largest protected areas on the planet and both are controlled by PNA countries.

@@

\subsection{Fishing effort markets and marine conservation}

We begin by simulating a spatial closure in a world where vessel-days can not be traded between countries. A spatial closure in Patch 1 always results in a loss or no-change in revenue from vessel-days, even when the stock moves freely between the protected and non-protected portions of the Patch (Fig. \ref{fig:PNA_model}A). The losses are caused by a drop in vessel-day prices for Patch 1, which results from a lower portion of the stock being available for harvest (Fig. \ref{fig:vessel_day_price_no_trading_plot}). Costs increase with reserve size ($R$), and decrease as within-patch movement ($\theta$) increases. Given the non-linear costs ($\beta > 1$), the closure-to-cost ratio can be greater than 1:1 if within-patch movement is low (\emph{i.e.} $\theta < 0.2$). The spatial closure in Patch 1 results in moderate increases in revenue for Patches 2 - 9 (Fig. \ref{fig:profits_PNA_notKIR_no_trading_plot}). When countries do not trade between them, the costs of conservation are large and incurred by Patch 1, but the benefits perceived by the eight other patches. Importantly, the increases in revenue to other patches do not compensate for the total losses to Patch 1.

The objective of a VDS is to allow for trading between parties and achieve the optimal distribution of effort in relation to fish distribution. In the case of the PNA, for example, spatial distribution and longitudinal shifts of tuna -particularly skipjack- have been linked to ENSO events \cite{lehodey_1997}. Vessels redistribute across PNA countries as they track fish, and tradable permits reduce costs to the parties \cite{aqorau_2018}. We explore the effect of a spatial closure in Patch 1, but now allow for vessel-days to be traded between Patches. Patch-level demand for vessel days is governed by the fishable stock size (\emph{i.e.} biomass not within the reserve). Trading occurs until vessel-day prices are the same across patches, given a total allowable effort (\emph{i.e.} 45,000 vessel-days) is reached. Here, costs from a closure in Patch 1 are significantly reduced by trading (Fig. \ref{fig:PNA_model}B-C). In this case, surplus of vessel-days in Patch 1 are sold to Patches 2 - 9. Vessel-day prices are constant across patches and lower than when there is no reserve, but still higher than when no trading is allowed (Fig. \ref{fig:vessel_day_price_with_trading_plot}). By optimally trading vessel-days, the cost to the implementing country can be reduced as much as 97\%.

Our previous analyses estimate loss in revenue from effort displacement. Annual vessel-days are often allocated based on a combination of historical within-patch effort and biomass. In the PNA, for example, sixty percent of the allocation is calculated based on EEZ effort over the last seven years and 40\% is calculated based on the 10-year average of each country’s share of estimated skipjack and yellowfin biomass within its EEZ. \footnote{This is explained in more detail in Article 12.5 of the 2012 Amendment to the Palau Agreement and in \cite{Hagrannsoknir2014}}. Trading vessel-days to other countries would imply that historical within-patch effort declines through time. The allocated days to the Patch with the spatial closure would eventually be reduced to the 40\% depending on biomass.

We explore the effect that allocation rules have through time when trading is allowed by projecting the fishery 50 years into the future with. We used a range of allocation rules that weight effort ($\alpha$) and biomass ($1 - \alpha$) differently to update annual allocations based on a seven-year running mean of these measures. Costs are greater when allocation is based only on patch-level historical effort (\emph{i.e.} $\alpha = 1$; Fig. \ref{fig:allocation_cost_plot}). However, costs can be significantly reduced when trading is allowed if allocation is based on historical biomass instead. When there is no trading, patch-level effort remains unchanged, and since any growth in stock is equally distributed across patches, allocation rules do not affect costs in a case without trading.

\subsection{Phoenix Islands Protected Area}

We use identification of fishing activity via Automatic Identification Systems (AIS) to track 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018 \cite{mccauley_2016,kroodsma_2018}. We track the spatial redistribution of the fleet and explore behavioral changes following the 2015 implementation of PIPA in Kiribati. We continuously observe 92 vessels for the 2012 - 2018 period. Of these, 64 vessels fished within PIPA at least once prior to its implementation, and we refer to them as the ``displaced vessels''. The remaining 28 vessels never fished in PIPA waters, and we refer to them as ``non-displaced vessels''. The group with the remaining 221 vessels contains vessels that were not continuously observed before and after the implementation of PIPA, and we refer to these as ``other vessels''. Our analyses incorporate the monthly NINO4 anomaly index (Fig. \ref{fig:nino_plot}) to control for the association that fish and vessels might have to environmental variation \cite{lehodey_1997,kroodsma_2018,aqorau_2018}.

A market for fishing effort and allocation rules can reduce the costs of displacing fishing effort, but a spatial closure is still likely to cause behavioral changes as vessels redistribute. We find that between 2015 and 2016, displaced vessels spent 3,916 and 2,249 less vessel-days in Kiribati and PNA waters, respectively (Figs. \ref{fig:all_PS_VDS_year}). Over the same period, non-displaced and other vessels spent 1,278 less days in Kiribati, but spent 9,853 more days in PNA waters overall. These result in a net loss of 5,195 vessel-days for Kiribati, and a net gain of 7,600 vessel-days at the PNA-level (\emph{i.e.} the other 8 countries). At a vessel-day price of \$9,000 USD per day, these would represent \$46.7 and \$68 million USD, respectively.

Annual vessel-days in Kiribati continued to decrease to just 7,479 in 2018, mainly due to displaced vessels allocating less time to Kiribati (Fig. \ref{fig:all_PS_VDS_year}B); Fig. \ref{fig:hist_kir_fishing}). Current effort in Kiribati waters are well below their 11,000 vessel-day allocation \cite{yeeting2018stabilising}. However, we observe a relatively constant 45,000 vessel-days at the PNA-level, suggesting that Kiribati is, in fact, able to trade vessel-days and reduce the costs of displacement. The spatial redistribution patterns of displaced vessels relative to non-displaced vessels are shown in Figure \ref{fig:fishing_raster_diff}.

We complement our analysis of change in observed vessel-days by looking at country-level data reported by the Pacific Islands Forum Fisheries Agency (FFA). We find that Kiribati's revenue went from \$148.8 million USD in 2015 to \$118.3 million USD in 2016, representing a decrease of \$30.5 million USD. Total PNA revenues showed a net increase of \$28 million USD (Fig. \ref{fig:total_PNA_revenues}). The largest decrease was observed for Kiribati, while the largest increase was also observed for Papua New Guinea.

Total fishing effort in the PNA increased between 2015 and 2016, but fishing area decreased due to PIPA implementation (Fig. \ref{fig:all_PS_VDS_year}). We inspect the crowding effects that may arise by applying more (or the same amount of) fishing effort over less fishing area with two indices of spatial overlap between displaced and non-displaced vessels: 1) the number of cells that had fishing activity from both groups for each month and 2) the correlation of presence/absence of fishing activity between both groups over one month. The two fleets significantly interact more with each other after the implementation of PIPA (Table \ref{tab:sp_corr} Fig. \ref{fig:sp_corr}). The number of cells with presence from both fleets and spatial correlation increase by a factor of four and three, respectively. Crowding might be driven by environmental variation concentrating fish at specific locations, but the NINO4 anomaly alone explained just 3\% and 7\% of the variation in our crowding measures, compared to the 70\% variation explained when accounting for PIPA implementaiton. We observe similar patterns when repeating the crowding exercise for Kiribati's EEZ only (Table \ref{tab:KIR_sp_corr}, Fig \ref{fig:KIR_sp_corr}).

Effort displacement and crowding are likely to induce behavioral changes in the fishing fleet. We calculate nine vessel-level measures that could capture responses to spatial closures: daily fishing hours, daily non-fishing at-sea hours, the proportion of fishing to non-fishing at-sea hours, daily distance traveled, daily mean distance from shore of fishing events (km), daily mean distance from port of fishing events (km), as well as monthly hours spent in PNA waters, Kiribati waters, and the High Seas (Fig. \ref{fig:all_panels}). We find no evidence of displaced vessels fishing more after PIPA implementation, and in fact observe a 27.5\% decrease ($p < 0.01$; Table \ref{tab:main_DID}) relative to non-displaced vessels. Likewise, we observe a 3.4\% decrease of fishing hours relative to total at-sea hours ($p < 0.01$; Table \ref{tab:main_DID}). Treated vessels traveled 23.3\% less distance, and fishing events occurred 32.9\% and 16.9\% closer to shore and to port, respectively. These changes in distance from shore and port are likely caused by redistribution, as we observe that displaced vessels fished 79.67\% and 49.6\% less in Kiribati and PNA waters respectively, compared to the trend observed for non-displaced vessels ($p < 0.01$). We do not observe a statistically significant increase in non-fishing at-sea hours (\emph{i.e.} a proxy for search time) and fishing hours on the High Seas. We repeat this analysis for sample restrictions where we exclude all Chinese vessels (Table \ref{tab:DID_without_CHN}), all PNA-owned vessels (Table \ref{tab:DID_without_PNA}) and all Taiwanese and USA vessels (Table \ref{tab:DID_without_USA_TWN}) and find qualitatively the same responses.

\subsection{Palau National Marine Sanctuary}

On October 28, 2015, the President of Palau signed into law the Palau National Marine Sanctuary (PNMS) Act. Starting in December 2020, this Act will close 500,000 km\textsuperscript{2} (about 80 percent of Palau’s EEZ) to commercial fishing activities, creating the 14th largest protected area in the world, effectively displacing all effort within its waters (Fig. \ref{fig:plw_2018} shows longline and purse seine fishing effort within PNMS for 2018). As part of the Nauru Agreement, Palau receives 700 purse seine vessel-days each year. They also  use a vessel-day scheme to manage the longline fishery  (10,500 vessel-days per year). Unlike the purse seine vessel days, longline vessel-days are not tradable within countries and are worth a fraction, at about \$200 USD each. Our previous analyses suggest that Palau should be able to lessen the costs of the closure by trading purse seine vessel-days (\$5.65 - \$8.75 million USD) with other PNA members, but might lose more than 50\% of the revenues from the longline vessel-days (\$2.1 million USD).
